---
title: "Flow TexRad DL Signatures Meta-analysis"
author: "P R Barber, M Tharmakulasingam"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggfortify)
library(msaWrapper)
library(pROC)
library(glmnet)
```

# Introduction

This follows on from TestSetExtraction.Rmd and the analyses that produce risk score signatures.

# Load Data

```{r}
flow_train <- read.csv("../train_flow.csv")
flow_test <- read.csv("../test_flow.csv")
texrad_train <- read.csv("../train_texrad.csv")
texrad_test <- read.csv("../test_texrad.csv")
dl_train <- read.csv("../train_dl.csv")
dl_test <- read.csv("../test_dl.csv")

data_namecv <- "flow_texrad_dl_sigs"

outcome_train <- read.csv("../train_outcomes.csv")
outcome_test <- read.csv("../test_outcomes.csv")

# Some rows are now missing from the outcomes since other cancers were excluded.
# AND we must be sure the sample order is the same throughout
# For all data types
merged_train <- merge(flow_train, outcome_train, by = "ID")
merged_test <- merge(flow_test, outcome_test, by = "ID")
flow_train <- subset(merged_train, select = -CancerDx)
flow_test <- subset(merged_test, select = -CancerDx)

merged_train <- merge(texrad_train, outcome_train, by = "ID")
merged_test <- merge(texrad_test, outcome_test, by = "ID")
texrad_train <- subset(merged_train, select = -CancerDx)
texrad_test <- subset(merged_test, select = -CancerDx)

merged_train <- merge(dl_train, outcome_train, by = "ID")
merged_test <- merge(dl_test, outcome_test, by = "ID")
dl_train <- subset(merged_train, select = -CancerDx)
dl_test <- subset(merged_test, select = -CancerDx)


flow_risk_train <- predictSPSignature(flow_train, "../flow/flow", Sps_regression_folder)
texrad_risk_train <- predictSPSignature(texrad_train, "../texrad/texrad", Sps_regression_folder)
dl_risk_train <- predictSPSignature(dl_train, "../dl_GAN32/dl", Sps_regression_folder)

train_set <- data.frame(ID = flow_train$ID,
                        flow_sig = flow_risk_train,
                        texrad_sig = texrad_risk_train,
                        dl_sig = dl_risk_train)

flow_risk_test <- predictSPSignature(flow_test, "../flow/flow", Sps_regression_folder)
texrad_risk_test <- predictSPSignature(texrad_test, "../texrad/texrad", Sps_regression_folder)
dl_risk_test <- predictSPSignature(dl_test, "../dl_GAN32/dl", Sps_regression_folder)

test_set <- data.frame(ID = flow_test$ID,
                        flow_sig = flow_risk_test,
                        texrad_sig = texrad_risk_test,
                        dl_sig = dl_risk_test)

groups <- c(1, 2, 3)
group_names <- c("flow", "texrad", "dl")

train_msa <- msaWrapperCreate(train_set, outcome_train$CancerDx, rowsLabelled = T, group = groups, groupLabels = group_names)
test_msa <- msaWrapperCreate(test_set, outcome_test$CancerDx, rowsLabelled = T, group = groups, groupLabels = group_names)

```

# Data Reduction

No need, only a few signature risk score covariates.

```{r}
#train_msa <- reduceByCorrelationWithinGroups(train_msa, threshold = 0.7)
saveRDS(train_msa, file = paste0(data_namecv, "_train_msa_reduced.Rds"))
```

```{r}
train_msa <- readRDS(paste0(data_namecv, "_train_msa_reduced.Rds"))
train_msa
train_msa$colLabels
```

# Train

Can take a long time, maybe run separately.

```{r}
if(!file.exists(paste0(data_namecv, "_SPS_model.Rds"))){
  
  SPS_model_cv <- buildSPSignature(train_msa, data_namecv, iterations = Sps_iterations,
                                   inifile = Sps_ini_file_template)
  
  saveRDS(SPS_model_cv, file = paste0(data_namecv, "_SPS_model.Rds"))
  file.rename("SPS_Signature_logfile.txt", paste0(data_namecv, "_SPS_Signature_logfile.txt"))
  
  # save train and test data in sps format for manual testing
  df <- train_msa$data
  df <- df[, order(colnames(df))] # order cols so train ans test match
  df$outcome = train_msa$outcome # add outcome, row order should be guaranteed
  export_SPS_file(df,
                  paste0("SPS_data_", data_namecv, "_train"),
                  type = 4,  # 4: ordinal class outcome data without sample IDs
                  C = 2)
  
  df <- test_msa$data[, colnames(test_msa$data) %in% colnames(train_msa$data)]
  df <- df[, order(colnames(df))] # order cols so train ans test match
  df$outcome = test_msa$outcome # add outcome, row order should be guaranteed
  export_SPS_file(df,
                  paste0("SPS_data_", data_namecv, "_test"),
                  type = 4,  # 4: ordinal class outcome data without sample IDs
                  C = 2)
  
  # ID = train_msa$rowLabels, but data reduction does not keep the row labels
}
```

```{r}
SPS_model <- readRDS(paste0(data_namecv, "_SPS_model.Rds"))
```

Check Train Performance

```{r}
data_name <- data_namecv
plotSPSignatureBetaSwimmers(SPS_model, show_best_performance_line = F, selectedSig = FALSE, textreducescale = 2)
plotSPSignatureOClass(SPS_model)

tiff(paste0("plotSPSignatureBetaSwimmers", ".tiff"), compression = "lzw", res = 600, width = 4000, height = 4000)
plotSPSignatureBetaSwimmers(SPS_model, show_best_performance_line = F, selectedSig = FALSE, textreducescale = 2)
dev.off()

tiff(paste0("plotSPSignatureOClass_train_box", ".tiff"), compression = "lzw", res = 600, width = 4000, height = 4000)
plotSPSignatureOClass(SPS_model)
dev.off()

```

```{r}
try(plotBetaTable(paste0(data_namecv, "/", Sps_regression_folder, "/"), cex = 0.7, rowspacing = 0.05, xmin = -2.5, betapos = 1.8, pvaluepos = 2.6,rankpos = 3.1,weightpos = 3.2, xmax = 3.2, useRank = TRUE))

tiff(paste0("plotBetaTable", ".tiff"), compression = "lzw", res = 600, width = 4000, height = 4000)
try(plotBetaTable(paste0(data_namecv, "/", Sps_regression_folder, "/"), cex = 0.7, rowspacing = 0.05, xmin = -2.5, betapos = 1.8, pvaluepos = 2.6,rankpos = 3.1,weightpos = 3.2, xmax = 3.2, useRank = TRUE))
dev.off()

```

## Get train Score and Signatures

```{r}
try(trainresultscv <- predictSPSignature(train_msa$data, data_namecv, Sps_regression_folder)) # will use trainresultscv as a flag for model exists

if(exists("trainresultscv")){
traingAUC <- getSPSignaturePerformance(trainresultscv, as.numeric(train_msa$outcome))

tiff(paste0("getSPSignaturePerformance_train_ROC", ".tiff"), compression = "lzw", res = 600, width = 4000, height = 4000)
getSPSignaturePerformance(trainresultscv, as.numeric(train_msa$outcome))
dev.off()

getSignatures(data_namecv, Sps_regression_folder)

getNormalisedSignatures(data_namecv, Sps_regression_folder)
}
```

```{r}
if(exists("trainresultscv")){
SPS_train_AUC_ci <- ci.auc(train_msa$outcome, trainresultscv)
cat(paste("Training AUC:", round(SPS_train_AUC_ci[2], digits = 2), "(", round(SPS_train_AUC_ci[1], digits = 2), "-", round(SPS_train_AUC_ci[3], digits = 2), ")"))
} else {SPS_train_AUC_ci = c(0, 0, 0)}
```

# Test set performance

## Get test set outcomes

```{r warning=FALSE}
  try(testresultscv <- predictSPSignature(test_msa$data, data_namecv, Sps_regression_folder))
  
if(exists("testresultscv")){

testAUC <- getSPSignaturePerformance(testresultscv, as.numeric(test_msa$outcome))

tiff(paste0("getSPSignaturePerformance_test_ROC", ".tiff"), compression = "lzw", res = 600, width = 4000, height = 4000)
getSPSignaturePerformance(testresultscv, as.numeric(test_msa$outcome))  # only makes tiff of the last plot
dev.off()

# get a tiff of the box plot
data <- data.frame(testresultscv, as.numeric(test_msa$outcome))
names(data) <- c("RiskScore", "Class")
data$Class <- as.factor(data$Class)
t.test_result <- t.test(RiskScore ~ Class, data = data)
p_value <- sapply(t.test_result$p.value, JNCI_pvals)
tiff(paste0("getSPSignaturePerformance_test_box", ".tiff"), compression = "lzw", res = 600, width = 4000, height = 4000)
ggplot(data, aes(x = Class, y = RiskScore, group = Class)) + 
        geom_boxplot(outlier.colour = NA) + geom_jitter(width = 0.2, 
        col = rgb(0.1, 0.2, 0.8, 0.3)) + theme_classic() + labs(title = paste("p-value", 
        p_value))
dev.off()
}
```

```{r}
if(exists("trainresultscv")){
SPS_test_AUC_ci <- ci.auc(test_msa$outcome, testresultscv)
cat(paste("Testing AUC:", round(SPS_test_AUC_ci[2], digits = 2), "(", round(SPS_test_AUC_ci[1], digits = 2), "-", round(SPS_test_AUC_ci[3], digits = 2), ")"))
} else {SPS_test_AUC_ci = c(0, 0, 0)}
```

# ElasNet Model (Lasso) Analysis

For this type of fusion we should get the signature predictions from the LASSO saved set.

```{r}
# This code is largely the same as above but now for lasso

flow_train <- read.csv("../train_flow.csv")
flow_test <- read.csv("../test_flow.csv")
texrad_train <- read.csv("../train_texrad.csv")
texrad_test <- read.csv("../test_texrad.csv")
dl_train <- read.csv("../train_dl.csv")
dl_test <- read.csv("../test_dl.csv")

data_namecv <- "flow_texrad_dl_sigs"

outcome_train <- read.csv("../train_outcomes.csv")
outcome_test <- read.csv("../test_outcomes.csv")


# Some rows are now missing from the outcomes since other cancers were excluded.
# AND we must be sure the sample order is the same throughout
# For all data types
merged_train <- merge(flow_train, outcome_train, by = "ID")
merged_test <- merge(flow_test, outcome_test, by = "ID")
flow_train <- subset(merged_train, select = -CancerDx)
flow_test <- subset(merged_test, select = -CancerDx)

merged_train <- merge(texrad_train, outcome_train, by = "ID")
merged_test <- merge(texrad_test, outcome_test, by = "ID")
texrad_train <- subset(merged_train, select = -CancerDx)
texrad_test <- subset(merged_test, select = -CancerDx)

merged_train <- merge(dl_train, outcome_train, by = "ID")
merged_test <- merge(dl_test, outcome_test, by = "ID")
dl_train <- subset(merged_train, select = -CancerDx)
dl_test <- subset(merged_test, select = -CancerDx)


# get predictions from saved models, try for strict result first
load("../flow/flow_ElsticNet_Lasso_model.rda")
if(!anyNA(ElasNet_model$strict_Predicted_risk) & sd(ElasNet_model$strict_Predicted_risk) > 0){
  flow_risk_train <- ElasNet_model$strict_Predicted_risk
  flow_risk_test <- as.vector(predictElasticNet(ElasNet_model,
                                                flow_test[,ElasNet_model$covar_names], 
                                                lambda = ElasNet_model$strict_lambda))
}else{
  flow_risk_train <- ElasNet_model$best_Predicted_risk
  flow_risk_test <- as.vector(predictElasticNet(ElasNet_model,
                                                flow_test[,ElasNet_model$covar_names], 
                                                lambda = ElasNet_model$best_lambda))
  print("WARNING - best model used for flow")
}


load("../texrad/texrad_ElsticNet_Lasso_model.rda")
if(!anyNA(ElasNet_model$strict_Predicted_risk) & sd(ElasNet_model$strict_Predicted_risk) > 0){
  texrad_risk_train <- ElasNet_model$strict_Predicted_risk
  texrad_risk_test <- as.vector(predictElasticNet(ElasNet_model, 
                                                  texrad_test[,ElasNet_model$covar_names],
                                                  lambda = ElasNet_model$strict_lambda))
}else{
  texrad_risk_train <- ElasNet_model$best_Predicted_risk
  texrad_risk_test <- as.vector(predictElasticNet(ElasNet_model, 
                                                  texrad_test[,ElasNet_model$covar_names],
                                                  lambda = ElasNet_model$best_lambda))
  print("WARNING - best model used for texrad")
}


load("../dl_GAN32/dl_ElsticNet_Lasso_model.rda")
if(!anyNA(ElasNet_model$strict_Predicted_risk) & sd(ElasNet_model$strict_Predicted_risk) > 0){
  dl_risk_train <- ElasNet_model$strict_Predicted_risk
  dl_risk_test <- as.vector(predictElasticNet(ElasNet_model, 
                                              dl_test[, ElasNet_model$covar_names], 
                                              lambda = ElasNet_model$strict_lambda))
}else{
  dl_risk_train <- ElasNet_model$best_Predicted_risk
  dl_risk_test <- as.vector(predictElasticNet(ElasNet_model, 
                                              dl_test[, ElasNet_model$covar_names], 
                                              lambda = ElasNet_model$best_lambda))
  print("WARNING - best model used for dl")
}

rm(ElasNet_model)  # for safety as saved ones have the same name as here

train_set <- data.frame(ID = flow_train$ID,
                        flow_sig = flow_risk_train,
                        texrad_sig = texrad_risk_train,
                        dl_sig = dl_risk_train)

test_set <- data.frame(ID = flow_test$ID,
                        flow_sig = flow_risk_test,
                        texrad_sig = texrad_risk_test,
                        dl_sig = dl_risk_test)

groups <- c(1, 2, 3)
group_names <- c("flow", "texrad", "dl")

train_msa <- msaWrapperCreate(train_set, outcome_train$CancerDx, rowsLabelled = T, group = groups, groupLabels = group_names)
test_msa <- msaWrapperCreate(test_set, outcome_test$CancerDx, rowsLabelled = T, group = groups, groupLabels = group_names)

```


Training the ElasticNet model. alpha=1 (lasso) is used as there are many features and lasso tends to perform better on data with larger features.

```{r warning=FALSE}
if(!file.exists(paste0(data_namecv, "_ElsticNet_Lasso_model.rda"))){
  ElasNet_model <-buildElasticNet(train_msa, iterations = Enet_iterations, alpha = 1)
  save(ElasNet_model, file=paste0(data_namecv, "_ElsticNet_Lasso_model.rda"))
}
```

```{r warning=FALSE}
Elasnetfilename <- paste0(data_namecv, "_ElsticNet_Lasso_model.rda")
load(Elasnetfilename)
plotElasticNetCoefs(ElasNet_model, xvar = "lambda", selectedSig=TRUE)
plotElasticNetCoefs(ElasNet_model, xvar = "lambda")
plotElasticNetPerformance(ElasNet_model)

print('Train with strict Lambda')
try(trainbestAUC <- getElasticNetPerformance(ElasNet_model$strict_Predicted_risk, as.numeric(train_msa$outcome)))

print('Sigantures- strict Lambda')
try(si <- getElasNetSignatures(ElasNet_model, is_optimal_signature=FALSE))
print(si)

try(strict_Predicted_response <- predictElasticNet(ElasNet_model, test_set[, train_msa$colLabels], 
                                             lambda = ElasNet_model$strict_lambda))
print('Test with strict Lambda')
try(testbestAUC <- getElasticNetPerformance(strict_Predicted_response, as.numeric(test_msa$outcome)))

```

```{r}
EN_train_AUC_ci <- ci.auc(train_msa$outcome, ElasNet_model$strict_Predicted_risk)
cat(paste("Training AUC:", round(EN_train_AUC_ci[2], digits = 2), "(", round(EN_train_AUC_ci[1], digits = 2), "-", round(EN_train_AUC_ci[3], digits = 2), ")"))

EN_test_AUC_ci <- ci.auc(test_msa$outcome, strict_Predicted_response)
cat(paste("Testing AUC:", round(EN_test_AUC_ci[2], digits = 2), "(", round(EN_test_AUC_ci[1], digits = 2), "-", round(EN_test_AUC_ci[3], digits = 2), ")"))
```

# Save AUCs

```{r}
auc_results <- data.frame(rbind(SPS_train_AUC_ci, SPS_test_AUC_ci, EN_train_AUC_ci, EN_test_AUC_ci))
colnames(auc_results) <- c("ci_lower", "AUC", "ci_upper")
write.csv(auc_results, file=paste0(data_namecv, "_AUC_results.csv"), row.names = T)
```
